---
title: ""
output:
  word_document:
    reference_docx: Sith_Dissertation_Style.docx
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# B. R Scripts used to run co-occurrence analysis and network construction
In order to construct the networks in this study, there are three functions:

1. compute correlation coefficient all pair of variables,

2. construct network from step1,

3. detect communities from step2.

Another function is used to visualize node properties of resulting networks to
determine the functions each node has in the network.

## Libraries

```{r libraries, eval=TRUE, suppress=message, eval=FALSE}

if (!require("igraph")) {
  install.packages("igraph", dependencies = TRUE)
}

if (!require("qgraph")) {
  install.packages("qgraph", dependencies = TRUE)
}

if(!require("ggplot2")) {
  install.packages("ggplot2", dependencies = TRUE)
}

if(!require("dplyr")) {
  install.packages("dplyr", dependencies = TRUE)
}

if(!require("RColorBrewer")) {
  install.packages("RColorBrewer", dependencies = TRUE)
}

```

## Step 1

The step in this analysis is to perform Spearman correlation with
each dataset. The computation starts at the first column containing injury
values. The output is a data frame with five columns for each dataset labeled
pair of injuries (2 columns), the correlation coefficient, *p*-value of the
pair.

```{r first_step, eval=FALSE}

cooc_table <- function(data){
  
  # create the object to store the result
  results <- matrix(nrow = 0, ncol = 6)
  
  for (a in 1:(length(names(data)) - 1)) {
    
    # every variable will be compared to every other variable, so there has
    # to be another loop that iterates down the rest of the columns
    
    for (b in (a + 1):length(names(data))) {
      
      # summing the abundances of species of the columns that will be
      # compared
      
      
      var1.sum <- sum(temp[, a], na.rm = TRUE)
      var2.sum <- sum(temp[, b], na.rm = TRUE)
      
      # if the column is all 0's no co-occurrence will be performed
      
      if (var1.sum > 1 & var2.sum > 1) {
        
        # compute correlation using Spearman's correlation measure
        # can change correlation methods based on one's objective
        test <- cor.test(data[, a], data[, b], method = "spearman",
                         na.action = "na.exclude", exact = FALSE)
        
        # There are warnings when setting exact = TRUE because of ties from the
        # output of Spearman's correlation
        # stackoverflow.com/questions/10711395/spear-man-correlation and ties
        # It would be still valid if the data is not normally distributed.
        
        # extract rho value
        rho <- test$estimate
        # extract p.value value
        p.value <- test$p.value
      }
      
      if (var1.sum <= 1 | var2.sum <= 1) {
        # if varible is too small, it will not be included in analysis
        rho <- 0
        p.value <- 1
      }
      
      # combine the value
      new.row <- c(names(data)[a], names(data)[b], rho, p.value, var1.sum,
                   var2.sum)
      
      # combine row of the result
      results <- rbind(results, new.row)
      
    }
    
  }
  # change the class to dataframe
  results <- as.data.frame(results)
  
  row.names(results) <- NULL
  
  # rename column
  names(results) <- c("var1", "var2", "rho", "p.value", "var1.sum", "var2.sum")
  
  # set class of rho data to be numeric
  results$rho <- as.numeric(as.character(results$rho))
  
  # set class of p.value to be numberic
  results$p.value <- as.numeric(as.character(results$p.value))

  return(results)
}

```

The second step is to construct a network of co-occurrence relationships in a
data frame format that can be used with the `igraph` package to create networks.
The input file contains the results from the step. The output from this is an
`igraph` object with all edges in the network containing the Spearman
correlation coefficient significant at *p*-values < 0.05.

```{r second_step, eval=FALSE}
plot_network <- function(data){
  
  #== adjust the vertices properties
  
  # creat a network from an edge list matrix
  
  net <- graph.edgelist(as.matrix(data[ ,c("var1","var2")]), directed = FALSE)
  
  # set color's node
  V(net)$color <- adjustcolor("khaki2", alpha.f = .8)
  # set color's frame node
  V(net)$frame.color <- adjustcolor("khaki2", alpha.f = .8)
  
  # set shape of node
  V(net)$shape <- "circle"
  
  # set size of node. 25 is suitable for showing grap on A4 paper
  V(net)$size <- 25
  
  V(net)$label.color <- "black"
  
  V(net)$label.font <- 2
  
  V(net)$label.family <- "Helvetica"
  
  V(net)$label.cex <- 1.0
  
  # == adjust the edge proterties
  
  # label weight of each node pair with correlation cofficient
  E(net)$weight <- as.matrix(table[, "rho"])
  
  E(net)$width <- abs(E(net)$weight)*10
  
  E(net)$color <- "steelblue2"
  
  # set graph layout to Fruchterman-Reingold layout
  net$layout <- layout_with_fr(net)
  
  # output is igraph object
  return(net)
}
```

The third step takes input from the second step and generates an `igraph`
object, which can be used with the `igraph` package to visualize the
network with community detection using the `optimal_cluster` function from the
`igraph` package.

```{r third_step, eval=FALSE}
cluster.network <- function(graph){
  
  # detect comminities with maximizing mudularity
  community <- cluster_optimal(graph, weights = abs(E(graph)$weight))
  
  # set colors for each community
  prettyColors <- brewer.pal(n = 8, name = 'Set2')
  
  # define color to the community
  V(graph)$color <- prettyColors[membership(community)]
  
  # set the size of netwrok
  V(graph)$size <- 15
  
  # set width of edge
  E(graph)$width <- abs(E(graph)$weight)*10
  
  return(graph)
}
```



The fourth and final step is the function, which is applied for extracting node
properties from each node in a network and plot those node properties with dot
graph with `ggplot2` package. The input file is an `igraph` object from the
second step. The output is a `ggplot2` graph showing node degree, betweenness,
clustering coefficient, clustering coefficient from the network. 

```{r fourth_step, eval=FALSE}
plot.node.centrality <- function(graph){
  
  # call function from qgraph
  cen <- centrality_auto(graph)$node.centrality
  cen$node <- row.names(cen)
  # compute clustering coeeficient from igraph package
  cen$CC <- igraph::transitivity(graph, type = "local", isolates = "zero")
  row.names(cen) <- NULL
  
  cus_theme <-   theme_bw() + theme(
    panel.grid.major.x =  element_blank(),
    panel.grid.minor.x =  element_blank(),
    panel.grid.major.y = element_line(color = "grey60", linetype = 3, size = 0.5),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 14))
  
  p1 <- cen %>% ggplot(aes(x= Degree, y = reorder(node, Degree))) +
    geom_point(size = 5, color ="red") +
    cus_theme + 
    xlab("Node degree") + 
    ylab("Node")    
  
  p2 <- cen %>% ggplot(aes(x= CC, y = reorder(node, CC))) + 
    geom_point(size = 5, color ="blue") +
    cus_theme +theme(axis.title.y = element_blank()) +
    xlab("Clustering Coef") 
  
  p3 <- cen %>% ggplot(aes(x= Betweenness, y = reorder(node, Betweenness))) + 
    geom_point(size = 5, color = "black") +
    cus_theme + theme(axis.title.y = element_blank()) +
    xlab("Betweenness")
  
  plot_grid(p1, p2, p3, labels=c("A", "B", "C"), ncol = 3, nrow = 1 )
  
}
```
