---
title: ""
output:
  word_document:
    reference_docx: Sith_Dissertation_Style.docx
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# C. Run differential network analysis in different seasons

In this appendix, the network correlation coefficients are visualized. Before
performing this function, two data sets from different conditions should be
selected to compare and construct a differerential network. Each data set is
needed to compute correaltion coefficient using the function named `cooc_table`
from Appendix B first. Then, the output from each data set will be able to
compare the differences using the `diff_cor` function.

**It is required to run Appendix B prior to this Appendix.**

# Libraries
```{R libraries, eval=FALSE}
if (!require("DiffCorr")) {
  install.packages("DiffCorr", dependencies = TRUE)
}

if (!require("igraph")) {
  install.packages("igraph", dependencies = TRUE)
}
```

# Differential season network
```{r differential_season_network, eval=FALSE}
# df_dry is the dataframe selected in dry season
# df_wet is the datafeam selected in wet season

# call function cooc_table defined in Appedix B

cor_dry <- cooc_table(df_dry) # dry season
cor_wet <- cooc_table(df_wet) # wet season

pair.list <- cor_dry
# = Dry season
cor_dry$rho[cor_dry$p.value > 0.05 ] <- 0
g1 <- graph.edgelist(as.matrix(cor_dry[1:2]), directed = FALSE)

# = Wet season
cor_wet2$rho[cor_df$p.value > 0.05 ] <- 0
g2 <- graph.edgelist(as.matrix(df_wet[1:2]), directed = FALSE)


# = Differential network
# g1 is network graph of dry season (condition1)
# g2 is network graph of wet season (condition2)

# create adjagency matrix from igraph object
adj.mat1 <- as.matrix(as_adjacency_matrix(g1, attr = "weight")) # dry season

adj.mat2 <- as.matrix(as_adjacency_matrix(g2, attr = "weight")) # wet season


# compare correlation coefficient of each pair of injuries of conditionA and 
# condition B

# call dfiff.corr function from 02_co_network_creation.R #

diff_comb <- diff.corr(adj.mat1 , adj.mat2)

# set the index that 
# if correlation coefficents of condiction A higher than B index value is positive
# if correlation coefficents of condiction A less than B index value is negative
diff_comb$index <- ifelse(abs(diff_comb$r1) - abs(diff_comb$r2) > 0, 1, -1)

# create differtial network in seasons

gdif <- graph.edgelist(as.matrix(diff_comb[1:2]), directed = FALSE)

E(gdif)$weight <- as.matrix(diff_comb$index)

V(gdif)$color <- adjustcolor("khaki2", alpha.f = .8)
V(gdif)$frame.color <- adjustcolor("khaki2", alpha.f = .8)
V(gdif)$shape <- "circle"
V(gdif)$size <- 25
V(gdif)$label.color <- "black"
V(gdif)$label.font <- 2
V(gdif)$label.family <- "Helvetica"
V(gdif)$label.cex <- 1.0
gdif$layout <- layout_with_fr(gdif)

# label edges following the result from comparison
# Dry season - Wet season 
# Dry > Wet > 0 ; labeled with organred4 color
# Dry < Wet < 0 ; labeled with royelblue4 color
E(gdif)[weight > 0]$color <- adjustcolor("orangered4" , alpha.f = .8)
E(gdif)[weight < 0]$color <- adjustcolor("royalblue4", alpha.f = .8)
```


# Using diff.corr function to compare differences in networks

This function performs Fisher’s *z*-test on the dataset from two different
conditions. The output is data frame with a column for the dataset label, each
injury pair, the correlation coefficient at condition A, *p*-value in the pair
at condition A, the correlation coefficient at condition B, *p*-value in the
pair at condition B, *p*-value of difference in the pair, and difference of two
correlation coefficients. The result will be used for constructing differential
network model with all edges in the network with differential co-occurrence
significant at *p*-values < 0.05, and labelled edges by color to identify the
condition that edges are represented; red edges represent differential
relationship in dry season, and blue edges represented differential relationship
in wet season.

```{r diff_cor, eval=FALSE}
diff.corr   <- function(data1, data2){
  
  ccc1 <- as.vector(data1[lower.tri(data1)])
  
  ccc2 <- as.vector(data2[lower.tri(data2)])
  
  n <- nrow(data1)
  
  N <- n * (n - 1)/2
  
  p1 <- rep(1, N)
  
  p2 <- rep(1, N)
  
  pdiff <- rep(1, N)
  
  diff <- rep(1, N)
  
  mol.names <- rownames(data1)
  
  p1 <- cor2.test(n1, ccc1)
  
  p2 <- cor2.test(n2, ccc2)
  
  pdiff <- compcorr(n1, ccc1, n2, ccc2)$pval
  diff <- ccc1 - ccc2
  
  pdiff[(is.na(pdiff)) == TRUE] <- 1
  
  myindex <- which((lower.tri(data1)) == TRUE, arr.ind = TRUE)
  
  mol.names1 <- mol.names[myindex[, 1]]
  
  mol.names2 <- mol.names[myindex[, 2]]
  
  fin.ind <- pdiff < 0.05
  
  # combine data
  res <- cbind(mol.names1[fin.ind], mol.names2[fin.ind], ccc1[fin.ind], p1[fin.ind], ccc2[fin.ind], p2[fin.ind], pdiff[fin.ind], diff[fin.ind])
  
  # correct format the data structure
  res <- as.data.frame(res)
  
  # name column
  names(res) <- c("var1", "var2", "r1", "p1", "r2", "p2", "p.difference", "difr")
  # set class of each variable
  res$var1 <- as.character(res$var1)
  res$var2 <- as.character(res$var2)
  res$r1 <- as.numeric(as.character(res$r1))
  res$p1 <- as.numeric(as.character(res$p1))
  res$r2 <- as.numeric(as.character(res$r2))
  res$p2  <- as.numeric(as.character(res$p2))
  res$p.difference <- as.numeric(as.character(res$p.difference))
  res$difr <- as.numeric(as.character(res$difr))
  
  return(res)
}
```


# Visualising differential networks
The last step visualises the differential netowrk analysis results. Correlation
coeffecients of two data sets (low yield data set and high yield data) are
computed using the function named `cooc_table` from Appendix B, first. Then, the
output from each data set will compared for differences using the `diff_cor`
function. The result is used for constructing differential network models with
all edges in the network with differential co-occurrence significant at 
*p*-values < 0.05, and select edges that are higher significantly co-occurrences
at lower yield levels.

```{r differential_yield_network, eval=FALSE}
cor_low <- cooc_table(df_low) # low yield level
cor_high <- cooc_table(df_high) # high yield level

# = low yield data
cor_low$rho[cor_dry$p.value > 0.05 ] <- 0
g1 <- graph.edgelist(as.matrix(cor_low[1:2]), directed = FALSE)
E(g1)$weight <- cor_low$rho

# = high yield data
cor_high$rho[cor_df$p.value > 0.05 ] <- 0
g2 <- graph.edgelist(as.matrix(df_high[1:2]), directed = FALSE)
E(g2)$weight <- cor_high$rho

# = Differential network
# g1 is network graph of low yield  (condition1)
# g2 is network graph of high yield (condition2)

# create adjagency matrix from igraph object
adj.mat1 <- as.matrix(as_adjacency_matrix(g1, attr = "weight")) # low yield

adj.mat2 <- as.matrix(as_adjacency_matrix(g2, attr = "weight")) # high yield

# compare correlation coefficient of each pair of injuries of conditionA and 
# condition B

# call dfiff.corr function from 02_co_network_creation.R #

diff_comb <- diff.corr(adj.mat1 , adj.mat2)

# set the index that 
# if correlation coefficents of condition A higher than B index value is positive
# if correlation coefficents of condition A less than B index value is negative
diff_comb$index <- ifelse(abs(diff_comb$r1) - abs(diff_comb$r2) > 0, 1, -1)

# create differtial network in yield levels

gdif <- graph.edgelist(as.matrix(diff_comb[1:2]), directed = FALSE)

E(gdif)$weight <- as.matrix(diff_comb$index)

V(gdif)$color <- adjustcolor("khaki2", alpha.f = .8)
V(gdif)$frame.color <- adjustcolor("khaki2", alpha.f = .8)
V(gdif)$shape <- "circle"
V(gdif)$size <- 25
V(gdif)$label.color <- "black"
V(gdif)$label.font <- 2
V(gdif)$label.family <- "Helvetica"
V(gdif)$label.cex <- 1.0
gdif$layout <- layout_with_fr(gdif)
E(gdif)[weight > 0]$color <- adjustcolor("grey60" , alpha.f = .8)

# this network will be considered pairs that higher significantly present at 
# low yield level because we assume that injury that highly present in low may
# cause yield loss

gdif <- delete.edges(gdif, which(E(gdif)$weight < 0))
```

